---
author: Debayan Sen (DSEN@bentley.edu)
---

Let's consider the function $f(x)=10x^2-16x+1$.
We focus not on the whole function, but just the expression on the right-hand side, $10x^2-16x+1$.


```{r}
formula <- expression( 10*x^2 - 16*x + 1 )
```

We can compute its derivative using the `D` function.

```{r}
D( formula, "x" )   # derivative with respect to x
```

R does not simplify the output, but if we do so ourselves, we find that $f'(x)=20x-16$.

If it had been a multi-variable function, we would need to specify
the variable with respect to which we wanted to compute a derivative.

```{r}
formula2 <- expression( x^2-y^2 )  # consider the formula x^2 - y^2
D( formula2, "y" )                 # differentiate with respect to y
```

That output says that $\frac{\partial}{\partial y}(x^2-y^2)=-2y$.
 
We can compute the second derivative by using the `D` function twice
and specifying the variables with respect to which we are computing the derivative.

```{r}
D( D( formula2, "x" ), "x" )  # second derivative with respect to x
```

```{r}
D( D( formula2, "x" ), "y" )  # mixed partial derivative
```
