---
author: Krtin Juneja (KJUNEJA@falcon.bentley.edu)
---

The solution below uses an example dataset about the teeth of 10 guinea pigs at three Vitamin C dosage levels (in mg) with two delivery methods (orange juice vs. ascorbic acid).  (See how to quickly load some sample data.)

```{r}
df <- ToothGrowth
```

To visually plot the means of the length of the tooth based on the Vitamin C dosage levels we can create a pointplot. We will use the `gplots` package. In the code below, `bars=TRUE` gives 95% confidence intervals for the means.

```{r}
# install.packages("gplots") # If you have not yet installed it
library(gplots)
plotmeans(len~dose, data=df, bars=TRUE)
```

The point plot informs us that as the dosage levels increase, the tooth length also increases. 

To obtain the actual numbers, we can use the code below.  The first line converts the numerical dosage values to a categorical variable, which may not be necessary if your data was already categorical.  

```{r}
df$dose.factor = as.factor(df$dose)
aov1 = aov(len~dose.factor, data=df) 
model.tables(aov1, type='means')
```

If you wish to display the difference between the overall mean and the group means, you can simply omit the `type='means'` parameter.

```{r}
model.tables(aov1)
```

To also see the specific values for the confidence intervals plotted earlier, we can use the `emmeans` package (Estimated Marginal Means or Least-Squares Means).

```{r}
# install.packages("emmeans") # If you have not yet installed it
library(emmeans)
emmeans(aov1,'dose.factor')
```
