---
author: Ni Shi (shi_ni@bentley.edu)
---

Let's create a small example DataFrame to use here (using the method from
how to create a data frame from scratch).
Naturally, you would apply this solution to your own data instead.

```{r}
df <- data.frame(
    Date  = c( '5/7/19', '5/10/19',   '5/11/19' ),
    Event = c(   'Work',   'Party', 'More work' )
)
df
```

We use the `as.Date()` function to convert a text column into dates.
If the input dates are not in the standard format, we can use the `format=` argument to change the format.
Note the difference between `%y` and `%Y`: The `%y` code means a 2-digit year, but the `%Y` code means a 4-digit year.

```{r}
df$Date = as.Date( df$Date, format='%m/%d/%y' )
df
```

It's often easier to handle date conversions while reading the data file.
You can use the `read_csv()` function in the `readr` package,
which will automatically recognize dates in some common formats.

Additionaly, you can use the `anytime()` function in the `anytime` package
to automatically parse strings as dates regardless of the format.

```{r}
# Use anytime() to attempt to parse various formats:
library(anytime)
examples <- c( "Nov 01 2022", "2022-11-01", "22-11-01" )
anytime( examples )
```

Note that it succeeded in two cases, but not the third.
