---
authors:
 - Krtin Juneja (kjuneja@falcon.bentley.edu)
 - Nathan Carter (ncarter@bentley.edu)
---

Usually, you have data you wish to compare, but we will use example data here.
We load the "oats" dataset from R's `MASS` package, about the yield of oats
from a split-plot field trial using three varieties (V) and four levels of
manurial treatment (N). The experiment was laid out in 6 blocks (B) of 3 main
plots, each split into 4 sub-plots. The varieties were applied to the main plots
and the manurial treatments to the sub-plots.

```{r}
# install.package('MASS')  # if you have not already done so, and want this data
library(MASS)
df <- oats
```

Before we perform the contrast test, let's verify that the yield of oats `Y`
depends on the nitrogen manurial treatment given to it `N`.

```{r}
aov1 <- aov(Y ~ N, data = df)
summary(aov1)
```

The $p$-value in the `Pr(>F)` column is below $\alpha=0.05$.
So at the 5% significance level, the yield differs according to the
nitrogen manurial treatment. We assume that the model assumptions are met but
do not verify them here.

We now want to perform a planned comparison test (or contrast test) on the data
to see whether there is a difference between the $N<0.5$ levels and the
$N>0.5$ levels. We will use the `fit.contrast` function in the `gmodels` package.
Since the order of the levels is 0, 0.2, 0.4 and 0.6, the
contrast coefficients will be $-0.5$, $-0.5$, $0.5$, $0.5$, respectively.

```{r}
# install.package('gmodels')  # if you have not already done so
library(gmodels)
fit.contrast(aov1, "N", coeff=c(-1/2,-1/2,1/2,1/2))
```

The $p$-value in the `Pr(>|t|)` column is below $\alpha=0.05$.
This tells us that there is a significant difference between the average yields
of the $N<0.5$ and $N>0.5$ levels.
