---
author: "Krtin Juneja (KJUNEJA@falcon.bentley.edu)"
---

When performing a linear regression, the following assumptions should be checked.

### 1. We have two or more columns of numerical data of the same length.

The solution below uses an example dataset about car design and fuel consumption from a 1974 Motor Trend magazine.  (See how to quickly load some sample data.)  We can see that our columns all have the same length.

```{r}
df <- mtcars
head(df)
```

### 2. Scatter plots we've made suggest a linear relationship.

Scatterplots are covererd in how to create basic plots, but after making the model,
we can also examine the residuals.

So let's make the model. Our predictors will be the number of cylinders and the weight of the car and the response will be miles per gallon. (See also how to fit a linear model to two columns of data.)

```{r}
model = lm(mpg~ cyl + wt, data=df)
```

We test for linearity with residual plots.  We show just one residual plot here; you should make one for each predictor.  R's plot function knows how to create residual plots.  (See also how to compute the residuals of a linear model.)

```{r}
plot(model, which = 1)
```

### 3. After making the model, the residuals seem normally distributed.

We can check this by constructing a QQ-plot, which compares the distribution of the residuals to a normal distribution.
Here we use SciPy, but there are other methods; see how to create a QQ-plot.

```{r}
plot(model, which = 2)
```

### 4. After making the model, the residuals seem homoscedastic.

This assumption is sometimes called "equal variance," and can be checked by the `regplot` function in Seaborn. We must first standardize the residuals, which we can do with NumPy.  We want to see a plot with no clear pattern; a cone shape to the data would indicate heteroscedasticity, the opposite of homoscedasticity.

```{r}
plot(model, which = 3) # assumption of equal variance
```
