---
author: "Nathan Carter (ncarter@bentley.edu) - replace with your name and email"
---

We first load a famous dataset, Fisher's irises, just to have some example
data to use in the code that follows.  (See how to quickly load some sample data.)

```{r}
library(datasets)
data(iris)
```

How big is the dataset?  The output shows number of rows then number of columns.

```{r}
dim(iris)  # Short for "dimensions."
```

What are the columns and their data types?  Can I see a sample of each column?

```{r}
str(iris)  # Short for "structure."
```

What do the first few rows look like?

```{r}
head(iris) # Gives 5 rows by default.  You can do head(iris,10), etc.
```

The easiest way to get summary statistics for an R `data.frame` is with the
`summary` function.

```{r}
summary(iris)
```

The columns from the original dataset are the column headings in the summary
output, and the statistics computed for each are listed below those headings.

We can also compute these statistics (and others) one at a time for any given
set of data points.  Here, we let `xs` be one column from the above
`data.frame` but you could use any vector or list.

```
xs <- iris$Sepal.Length

mean( xs )           # mean, or average, or center of mass
median( xs )         # 50th percentile
quantile( xs, 0.25 ) # compute any percentile, such as the 25th
var( xs )            # variance
sd( xs )             # standard deviation, the square root of the variance
sort( xs )           # data in increasing order
sum( xs )            # sum, or total
```
